Distributional Trees and Forests

### 2017-12-05 pkg/disttree/inst/CMStatistics2017/slides_rain.Rnw
edited application section and references
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1813 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 252f6545fa6410ba6db6141f20a5ec1c441237cd

\title{Distributional Trees and Forests}
\author{Lisa Schlosser, Torsten Hothorn,\\ Heidi Seibold, Reto Stauffer, Achim Zeileis}
\setbeamerfont{url}{size*={11.5pt}{13pt},series=\mdseries}
\setbeamerfont{footnote}{size=\tiny}
\renewcommand\footnoterule{}
\URL{https://R-Forge.R-project.org/projects/partykit/}
\headerimage{3}

@@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace*{0.5cm}
%Task:
\begin{itemize}
\item Goal: Predict total precipitation amount.
\item Input: Numerical Weather Prediction (NWP).
\item Problem: Uncertain initial conditions and \\
\hspace{1.7cm}unknown processes.
\item Idea: Statistical model with numerical predictions as \\
\hspace{1cm}covariates.
\item Distribution Assumption:\\
\vspace{0.2cm}
\hspace{1cm}$(\text{Total precipitation})^\frac{1}{1.6} \sim \textit{cens}\mathcal{N}(\mu,\sigma^2)$
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace*{0.2cm}
Data:
\begin{itemize}
\item 95 observation stations in Tyrol.
\item Daily 24h precipitation sums from July of 28 years (1985--2012)\footnote{provided by the National Hydrographical Service (EHYD)}.
\item Numerical ensemble predictions of meteorological quantities\footnote{provided by the {U.S.} National Oceanic and Atmospheric Administration (NOAA)\\
\hspace{0.32cm}using the Global Ensemble Forecast System (GEFS)}: \\
\hspace{0.5cm}11 members each using slightly different \\
\hspace{0.5cm}perturbed initial conditions.
\end{itemize}

@@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace*{0.2cm}
Variable Selection:
\begin{itemize}
\item Complex terrain $\rightarrow$ unknown influences.
\item 82 Covariates deduced from 14 base variables:
\begin{itemize}
\item Total precipitation
\item Convective available potential energy
\item Downwards short wave radiation flux (``sunshine'')
\item Mean sea level pressure 
\item Mean sea level pressure difference
\item Preciptable water
\item 2m maximum temperature
\item Total column-integrated condensate
\item Temperature 
\item Temperature differences in altitude
\end{itemize}
\item Variations such as ensemble minimum/maximum/mean/standard deviation.
\end{itemize}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace*{0.2cm}
Application for one station:
\begin{itemize}
\item Learn forest model on data from 24 years.
\item Evaluate on 4 years.
\item 10 times 7-fold cross validation.
\item Compare to commonly used heteroscedastic censored Gaussian models:\\
\begin{itemize}
\item Prespecified GAM: variable selection based on meteorological expert knowledge.
\item Boosted GAM: automatic variable selection.
\item Ensemble Model Output Statistics (EMOS): linear predictors using only total precipitation predictions as covariates.
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace*{0.2cm}
Application for all 95 stations:
\begin{itemize}
\item Learn forest model on data from 24 years (1985--2008).
\item Evaluate on 4 years (2009--2012).
\item Compare to commonly used heteroscedastic censored Gaussian models:\\
\begin{itemize}
\item Prespecified GAM: variable selection based on meteorological expert knowledge.
\item Boosted GAM: automatic variable selection.
\item Ensemble Model Output Statistics (EMOS): linear predictors using only total precipitation predictions as covariates.
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{References (1)}

Hothorn T, Hornik K, Zeileis A (2006).
 \dquote{Unbiased Recursive Partitioning: A Conditional Inference Framework.}
 \textit{Journal of Computational and Graphical Statistics},
 \textbf{15}(3), 651--674.
 \doi{10.1198/106186006X133933}

@@

@@

Seibold H, Zeileis A, Hothorn T (2017).
  \dquote{Individual Treatment Effect Prediction for Amyotrophic Lateral Sclerosis Patients.}
  \textit{Statistical Methods in Medical Research}, 
  \textbf{12}(1), 45--63.
  \doi{10.1177/0962280217693034}

@@

Hothorn T, Zeileis A (2017).
  \dquote{Transformation Forests.}
  \emph{arXiv 1701.02110}, arXiv.org E-Print Archive.
  \url{http://arxiv.org/abs/1701.02110}

@@

@@

% Reference for the GFS reforecast version 2 data set 
Hamill T M, Bates G T, Whitaker J S, Murray D R, Fiorino M, Galarneau T J Jr., Zhu Y, Lapenta W (2013).
  \dquote{{NOAA}'s Second-Generation Global Medium-Range Ensemble Reforecast Dataset}
  \textit{Bulletin of the American Meteorological Society}, 
  \textbf{94}(10), 1553--1565.
  \doi = {10.1175/BAMS-D-12-00014.1}

@@

Messner J W,  Mayr G J, Zeileis A (2016).
  \dquote{Heteroscedastic Censored and Truncated Regression with {crch}}
  \textit{The R Journal}, 
  \textbf{8}(1), 173--181.
  \url{https://journal.R-project.org/archive/2016-1/messner-mayr-zeileis.pdf}

@@

Breiman L (2001).
  \dquote{Random {Forests}}
  \textit{Machine Learning}, 
  \textbf{45}(1), 5--32.
  \doi{10.1023/A:1010933404324}

\begin{frame}
\frametitle{References (3)}

\footnotesize
Gneiting T, Raftery A E, Westveld III A H, Goldman T (2005).
  \dquote{Calibrated probabilistic forecasting using ensemble model output statistics and minimum CRPS estimation}
  \textit{Monthly Weather Review}, 
  \textbf{133}(5), 1098--1118.
  \doi{https://doi.org/10.1175/MWR2904.1}

@@

{CGIAR-CSI} (2016).
  \dquote{{SRTM} 90m Digital Elevation Database {v4.1}}
  \text{Available at \url{http://srtm.csi.cgiar.org}}, 
  \text{Accessed: {J}uly 10 2016}.

>>>>

\title{Distributional Trees and Forests}
\author{Lisa Schlosser, Torsten Hothorn,\\ Heidi Seibold, Reto Stauffer, Achim Zeileis}
\setbeamerfont{url}{size*={11.5pt}{13pt},series=\mdseries}
%\renewcommand{\mysize}{\fontsize{7.7pt}{9pt}\selectfont}
\URL{https://R-Forge.R-project.org/projects/partykit/}
\headerimage{3}

@@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{0.4cm}
\textbf{Goal:} Predict total precipitation amount in alpine regions.\\
\medskip
\textbf{Input:} Numerical Weather Prediction (NWP):\\ 
Global physical model numerically simulating atmospheric processes on $50\times 50 \text{km}^2$ grid.\\
\medskip
\textbf{Problem:} Uncertain initial conditions, unresolved processes.\\ %$\rightarrow$ requires post-processing.\\
\vspace{1.3cm}
%\bigskip
\textbf{Distribution Assumption:}
%\vspace{0.1cm}
%\hspace{1cm}
$(\text{Total precipitation})^\frac{1}{1.6} \sim \textit{c}\mathcal{N}(\mu,\sigma^2)$\\
\bigskip
\textbf{Idea:} Statistical model with numerical predictions as covariates.
\end{frame}

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{0.4cm}
\textbf{Observation data} (National Hydrographical Service):\\
\begin{itemize}
\item 95 observation stations in Tyrol, Austria.\\
\item Daily 24h precipitation sums from July of 28 years (1985--2012).
\end{itemize}
\vspace{1cm}
%\bigskip
\textbf{NWP model output} (Global Ensemble Forecast System):\\
\begin{itemize}
\item Numerical ensemble predictions of various meteorological quantities.\\
\item 11 members each using slightly different perturbed initial conditions.
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace*{0.2cm}
\textbf{Base Variables:}
\begin{itemize}
\item Total precipitation
\item Convective available potential energy
\item Downwards short wave radiation flux (``sunshine'')
\item Mean sea level pressure
\item Preciptable water
\item 2m maximum temperature
\item Total column-integrated condensate
\item Temperature 
\item Temperature differences in altitude
\end{itemize}
\medskip
\textbf{Variations:} Ensemble min/max/mean/standard deviation\\
\vspace{0.1cm}
$\rightarrow$ 82 covariates.

@@

@@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{0.2cm}
Application for one station:
\begin{itemize}
\item Learn forest model on data from 24 years.
\item Evaluate on 4 years.
\item 10 times 7-fold cross validation.
\item Compare to commonly used heteroscedastic censored Gaussian models:
\end{itemize}
%\begin{itemize}
%\item Prespecified GAM: variable selection based on meteorological expert knowledge.
%\item Boosted GAM: automatic variable selection.
%\item Ensemble Model Output Statistics (EMOS): linear predictors using only total precipitation predictions as covariates.
%\end{itemize}
\vspace{0.1cm}
\begin{table}[t!]
\footnotesize
\nodindent
\hskip-0.4cm\begin{tabular}{ l l l }
\hline
Model & Type & Variable Selection \\
\hline
\vspace{-0.15cm}
\textbf{Forest} & recursive    & automatic \\ 
                & partitioning &           \\ 
\hline
\vspace{-0.15cm}
\textbf{prespecified GAM} & spline  & based on expert \\
                          & in each & knowledge      \\
\hline
\vspace{-0.15cm}
\textbf{boosted GAM} & spline  & automatic  \\
                     & in each &            \\
\hline
\vspace{-0.15cm}
\textbf{EMOS} & linear & total precipitation mean \\
              &        & and standard deviation   \\
\hline
\end{tabular}
\end{table}

@@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{0.2cm}
Application for all 95 stations:
\begin{itemize}
\item Learn forest model on data from 24 years (1985--2008).
\item Evaluate on 4 years (2009--2012).
\item Compare to commonly used heteroscedastic censored Gaussian models:\\
\end{itemize}
%\begin{itemize}
%\item Prespecified GAM: variable selection based on meteorological expert knowledge.
%\item Boosted GAM: automatic variable selection.
%\item Ensemble Model Output Statistics (EMOS): linear predictors using only total precipitation predictions as covariates.
%\end{itemize}
\vspace{0.1cm}
\begin{table}[t!]
\footnotesize
\nodindent
\hskip-0.4cm\begin{tabular}{ l l l }
\hline
Model & Type & Variable Selection \\
\hline
\vspace{-0.15cm}
\textbf{Forest} & recursive    & automatic \\ 
                & partitioning &           \\ 
\hline
\vspace{-0.15cm}
\textbf{prespecified GAM} & spline  & based on expert \\
                          & in each & knowledge      \\
\hline
\vspace{-0.15cm}
\textbf{boosted GAM} & spline  & automatic  \\
                     & in each &            \\
\hline
\vspace{-0.15cm}
\textbf{EMOS} & linear & total precipitation mean \\
              &        & and standard deviation   \\
\hline
\end{tabular}
\end{table}
\end{frame}

@@

\begin{frame}
\frametitle{References}

\vspace{-0.2cm}

\footnotesize
Hothorn T, Hornik K, Zeileis A (2006).
 \dquote{Unbiased Recursive Partitioning: A Conditional Inference Framework.}
 \textit{Journal of Computational and Graphical Statistics},
 \textbf{15}(3), 651--674.
 \doi{10.1198/106186006X133933}

@@

%\medskip

%Seibold H, Zeileis A, Hothorn T (2017).
%  \dquote{Individual Treatment Effect Prediction for Amyotrophic Lateral Sclerosis Patients.}
%  \textit{Statistical Methods in Medical Research}, 
%  \textbf{12}(1), 45--63.
%  \doi{10.1177/0962280217693034}

%\medskip

%Hothorn T, Zeileis A (2017).
%  \dquote{Transformation Forests.}
%  \emph{arXiv 1701.02110}, arXiv.org E-Print Archive.
%  \url{http://arxiv.org/abs/1701.02110}

@@

Gneiting T, Raftery A E, Westveld III A H, Goldman T (2005).
  \dquote{Calibrated probabilistic forecasting using ensemble model output statistics and minimum CRPS estimation}
  \textit{Monthly Weather Review}, 
  \textbf{133}(5), 1098--1118.
  \doi{10.1175/MWR2904.1}

@@

@@

\vspace{-0.2cm}

@@

% Reference for the GFS reforecast version 2 data set 
Hamill T M, Bates G T, Whitaker J S, Murray D R, Fiorino M, Galarneau T J Jr., Zhu Y, Lapenta W (2013).
  \dquote{{NOAA}'s Second-Generation Global Medium-Range Ensemble Reforecast Dataset}
  \textit{Bulletin of the American Meteorological Society}, 
  \textbf{94}(10), 1553--1565.
  \doi{10.1175/BAMS-D-12-00014.1}

@@

%Breiman L (2001).
%  \dquote{Random {Forests}}
%  \textit{Machine Learning}, 
%  \textbf{45}(1), 5--32.
%  \doi{10.1023/A:1010933404324}

Messner J W,  Mayr G J, Zeileis A (2016).
  \dquote{Heteroscedastic Censored and Truncated Regression with {crch}}
  \textit{The R Journal}, 
  \textbf{8}(1), 173--181.\\
  \scriptsize
  \url{https://journal.R-project.org/archive/2016-1/messner-mayr-zeileis.pdf}
  \footnotesize

@@

@@

%\medskip

%{CGIAR-CSI} (2016).
%  \dquote{{SRTM} 90m Digital Elevation Database {v4.1}}
%  \text{Available at \url{http://srtm.csi.cgiar.org}}, 
%  \text{Accessed: {J}uly 10 2016}.
